USAMS Hybrid Source Summary

### 2021-01-24 hgis_summary.Rmd
Update summary.

Best currents for the source are around 10uA 12C- or roughly 1/10 of graphite
currents. This is in line with what other hybrid sources produce.  The source
typically produces best current at around 10uLs-1 CO2. Literature values give
1.5uLs-1 as optimal, but we've never seen that, and pressures, capillaries and
flows in at least one of these papers doesn't make sense.

Initially, precision on modern targets was thought to be as good as 1% (SD of
multiple measurements of a modern target) with directly introduced bottle gas,
which is in agreement with literature values. Revisiting these data show
precision in the range of 2-4%. Typical precision in tests of inlet systems has
been in the range of 2-4%.

Measured value of tank modern gas agrees with the accepted value within error
when normalizing to solid graphite standards.  The system seems to have a
constant blank contribution from the target and source of about 100nA 12C- or 1%
of a sample producing 10uA 12C-. Helium does not seem to add significantly to
this blank. Since the blank current is constant, the blank contribution is
inversely related to the sample current, ie. if the sample current is 5uA, 2% of
that will be from the blank and so on.

The above measurements are at optimum sample current of 6-10uA 12C-. Precision
and measured ratio both decrease as sample current drops below the optimum
range. The ratio effects may be related to the "constant current" blank
discussed above.

@@

The open split interface supplies a helium-co2 mixture to a split where a
capillary delivers ~10uL/min to the source and the rest is vented. The current
split starts showing "breakthrough" of lab air at around 150uL/min flow.
Currently using 3m of 50um capillary with a bulkhead connector at the source
cage.  Precision of measurements of modern CO2 has been 2% at best, 10% at
worst, with better precision most recently. This is much less precise than the
expected 1%.  Accuracy is reasonable within the large errors on gas
measurements. Uncorrected measurements of dead CO2 range from 1% to 5%
depending on sample current.

@@

First tests of displacing pure CO2 from a vial into the open split using helium
are promising. Helium flowing at 100uLm-1 produces current from 7mL of CO2 for
over 2 hours. Currents do decrease with time due to dilution, but by supplying
higher than optimal CO2 flow to the source initially, stable currents can be
maintained for some time.

@@

Q: Why is the precision not as expected? Can we replicate the
literature/initial test precision?

A: Unknown. Tests with the best performing conditions (bottle gas through
capillary to source) give 2% precision, and revisiting early test data shows
precision of 2%. Precision is driven by variability between measurements of a
sample. Investigating correlations here may help. We should also look at
variability between identical samples to get a sense of reproducibility.

## Development
2021-01-15 - Pure CO2 and displacement tests using vial and double needle. Tested precision with vial setup by flowing CO2 at 200kPa (high flow, roughly 400ul/min) into open split. Followed this by switching to helium flow to vial at 100kPa (200ul/min). Repeated test with a second vial. Built second dual needle setup for use in vial filling (can also use Gilson).

>>>>

Best currents for the source are around 10uA 12C- or roughly 1/10 of graphite
currents. This is in line with what other hybrid sources produce.  The source
typically produces best current at around 5uLs-1 CO2. Literature values give
1.5uLs-1 as optimal, but we've never seen that, and pressures, capillaries and
flows in at least one of these papers doesn't make sense.

Precision taken as the standard error of the mean of repeated runs of a sample is around 1%. Precision is typically a little worse than predicted by counting statistics. More work is needed to look at whether this reflects variability between samples.

Measured value of tank modern gas normalized to solid graphite standards ranges from agreeing with solid OX-I to 7% lower than expected for low current samples. Normalizing to tank CO2 is probably the best method to account for variability in hybrid source performance.

The system seems to have a constant blank contribution from the target and
source of about 100nA 12C- or 1% of a sample producing 10uA 12C-. Helium does
not seem to add significantly to this blank. Since the blank current is
constant, the blank contribution is inversely related to the sample current, ie.
if the sample current is 5uA, 2% of that will be from the blank and so on.

The above measurements are at optimum sample current of 6-10uA 12C-. Precision
and measured ratio both decrease as sample current drops below the optimum
range. The ratio effects may be related to the "constant current" blank
discussed above. More work is needed to see whether this is just a blank effect or whether there are other factors at low current.

@@

The open split interface supplies a helium-co2 mixture to a split where a
capillary delivers ~10uL/min to the source and the rest is vented. The current
split starts showing "breakthrough" of lab air at around 150uL/min flow. A new open split with lower volume should have improved this, but testing is needed.

Currently, 250cm of 50um capillary defines flow rate from the open split into
the source, with a bulkhead connector at the source cage.  Precision of
measurements of modern CO2 has been better than 1% (std error of mean of
measurements of a target) for high current samples, with precision dropping off
as current decreases below 6uA or so. Accuracy is reasonable within the large
errors on gas measurements. Uncorrected measurements of dead CO2 range from 1%
to 5% of modern depending on sample current.

@@

First tests of displacing pure CO2 from a vial into the open split using helium
are promising. Helium flowing at 100uLm-1 produces current from 7mL of CO2 for
over 2 hours. Currents do decrease with time due to dilution, but by supplying
higher than optimal CO2 flow to the source initially, stable currents can be
maintained for some time. Targets have shown "pulses" of higher current. Not sure why this is, but I suspect it's either buildup of graphite or other available carbon on the target, or target switching to the "high state" as sometimes happens with solid targets.

@@

Q: Does standard error of the mean as a sample error agree with between target variability?

A: Unknown. We should look at variability between identical samples to get a sense of reproducibility. Some data are available, but no controlled replicates have been run.

## Development

2021-01-22 - Ran pure CO2 at different pressures to look at effect of current on precision and ratio. Repeated test for live and dead tank CO2.

2021-01-15 - Pure CO2 and displacement tests using vial and double needle. Tested precision with vial setup by flowing CO2 at 200kPa (high flow, roughly 400ul/min) into open split. Followed this by switching to helium flow to vial at 100kPa (200ul/min). Repeated test with a second vial. Built second dual needle setup for use in vial filling (can also use Gilson).
